Tree size, (microhabitat)?, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-08-15 manuscript/climate_sensitivity_draft.Rmd
Update climate_sensitivity_draft.Rmd

**Table 1. Summary of hypotheses, predictions and results** 
![Table 1](tables_figures/Table1.png){width=500px}

@@

*Climate and drought years*
[**add description of climate data used in Fig. 1, NEON vertical profiles**]

We identified droughts within the time period as 1950-2009, defining drought as events where where peak growing season climatic conditions were anomalously dry relative to the long-term record and where tree growth was substantially reduced for at least some components of the community. To identify years with widespread reductions in tree growth, we used the pointRes package [@R-pointRes] in R (version 3.5.3) to determine drought periods based on trees' drought resistance, which is defined as the ratio between the performance during and before the disturbance [@lloret_components_2011]. Candidate drought years were defined if >50% of the cored trees experienced <30% growth in a year compared to the previous 5 years for at least three species-canopy class combinations, where canopy class included canopy (dominant and co-domiant) and subcanopy (suppressed and intermediate). **[this includes 1991- why did we drop it? [issue #52](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/52)]** Candidate droughts were then cross-validated with the regional Palmer Drought Severity Index (PDSI) values [**source?**] **(and MJJA PET-PRE (droughts identified in Helcoski)?)**. **[This paragraph needs work. See issue #52](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/52). Put climate criteria first.]**

@@

*Descriptions of Droughts*
In the 60-year period between 1950 and 2009, there were three droughts that met our definition of anomalously dry climatic conditions coupled with substantial reductions in tree growth for at least some portions of the community: 1964-1966, 1977, and 1999 (Fig. 1). We note that 1964 and 1966 were separately identified as drought years and grouped into a single multi-year drought period because 1965 was also anomalously hot and dry, albiet not identified as a pointer year. Climatically, these droughts included four of the five years between 1950 and 2009 with greatest moisture deficit (PET-PRE) during the peak growing season months of May-August, which are the months to which annual tree growth at this site is most sensitive [@helcoski_growing_2019]. Specifically, 1964, 1966, 1977, and 1999 had mean MJJA PET-PRE of 83.87, 83.37, 86.97, and 80 mm mo<sup>-1</sup>, respectively. The year 2007 also ranked in the top five driest (PET-PRE = 82.13 mm mo<sup>-1</sup>), but it did not have particualarly low PDSI and was not identified as a pointer year. **add stats for PDSI, T, precip**

- physical characteristics of drought: duration, min PDSI (and when occurred), see [@slette_how_2019] to see how they suggest reporting each one
- sum nb.series for each scenario, mult by perc.neg (pointers df), define thresholds (e.g. 75% of all trees experienced the growth reduction of 40% or more)

@@

**Figure 1. Time series of peak growing season (May-August) climate conditions and residual chronologies for each species.** Droughts analyzed here are indicated by dashed lines, and shading indicates the pre-drought period used in calculations of the resistance metric. Figure modified from [@helcoski_growing_2019].

>>>>

**Table 1. Summary of hypotheses, predictions and results**

![Table 1](tables_figures/Table1.png){width=500px}

@@

*Identification of drought years*
[**add description of climate data used in Fig. 1, NEON vertical profiles**]

We identified droughts within the time period 1950-2009, defining drought [@slette_how_2019] as events where peak growing season climatic conditions were anomalously dry relative to the long-term record and where tree growth was substantially reduced for at least some components of the community. To identify years with widespread reductions in tree growth, we used the pointRes package [@R-pointRes] in R (version 3.5.3) to determine drought periods based on trees' drought resistance, which is defined as the ratio between the performance during and before the disturbance [@lloret_components_2011]. Candidate drought years were defined if >50% of the cored trees experienced <30% growth in a year compared to the previous 5 years for at least three species-canopy class combinations, where canopy class included canopy (dominant and co-domiant) and subcanopy (suppressed and intermediate). **[this includes 1991- why did we drop it? [issue #52](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/52)]** Candidate droughts were then cross-validated with the regional Palmer Drought Severity Index (PDSI) values [**source?**] **(and MJJA PET-PRE (droughts identified in Helcoski)?)**. **[This paragraph needs work. See issue #52](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/52). Put climate criteria first.]**

[averaged resistance values for 1964-1966]

@@

*Descriptions of Droughts*
In the 60-year period between 1950 and 2009, there were three droughts that met our criteria of anomalously dry climatic conditions coupled with substantial reductions in tree growth for at least some portions of the community: 1964-1966, 1977, and 1999 (Fig. 1). The years 1964 and 1966 were separately identified as drought years and grouped into a single multi-year drought period because 1965 was also anomalously hot and dry, albiet not identified as a pointer year, and because inclusion of 1964 in the reference period for 1966 would have biased analyses. Climatically, these droughts included four of the five years between 1950 and 2009 with greatest moisture deficit (PET-PRE) during the peak growing season months of May-August, which are the months to which annual tree growth at this site is most sensitive [@helcoski_growing_2019]. Specifically, 1964, 1966, 1977, and 1999 had mean MJJA PET-PRE of 83.87, 83.37, 86.97, and 80 mm mo<sup>-1</sup>, respectively. The year 2007 also ranked in the top five driest (PET-PRE = 82.13 mm mo<sup>-1</sup>), but it did not have *particualarly low* PDSI and was not identified as a pointer year. **add stats for PDSI, T, precip**

Community-level tree growth responses to these droughts were modest, with modal resistance values of #, #, and # for 1964-1966, 1977, and 1999, respectively (Fig. 1b). In each drought, roughly 30% of the cored trees suffered growth reductions of 30% or more (resistance <= 0.7): *28.2%* in 1964-66, 31.2% un 1977, and 26.2% in 1999. *Some* trees exhibited increased growth: (resistance > 1.0): #% in 1964-66, #% un 1977, and #% in 1999.

@@

**Figure 1. ** (a) Time series of peak growing season (May-August) climate conditions and residual chronologies for each species. Droughts analyzed here are indicated by dashed lines, and shading indicates the pre-drought period used in calculations of the resistance metric. Figure modified from [@helcoski_growing_2019]. (b) density plots of community- wide resitance values for each drought.
